---
title: Intro to Networks
author: Maya Gans
date: '2019-06-12'
slug: []
categories:
  - Network Science
tags:
  - Networks
  - iGraph
subtitle: ''
summary: ''
authors: []
lastmod: '2019-06-12T14:42:18-06:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

## Chapter 2 

# PIVOTAL NODE

```{r, include=FALSE}
library(igraph)
library(ggraph)
library(gridExtra)
```

Create a simple, cyclic graph
```{r}
g1 <- graph( c("E", "A", "A", "B", "B", "D", "D", "C", "C", "F", "F", "E", "B", "C"), directed = FALSE)
```

When the edge list has vertex names, the number of nodes is not needed

```{r}
ggraph(g1, layout = "kk") + 
  geom_edge_link(colour = "lightgrey") + 
  geom_node_point(colour = "grey", size = 7) +
  geom_node_text(label = V(g1)$name) +
  theme_graph() +
  theme(legend.position="none")
```

### Node **B** is pivotal for the pair consisting of A and C and A and D. 
```{r}
# add edge data for pivotal nodes
E(g1)$pivotalB <- c("No", "Yes", "Yes", "No","No", "No", "Yes")
V(g1)$pivotalB <- c("No","No","Yes","No","No","No")

ggraph(g1, layout = "kk") + 
  geom_edge_link(aes(colour = E(g1)$pivotalB)) + 
  geom_node_point(aes(colour = V(g1)$pivotalB), size = 7) +
  geom_node_text(label = V(g1)$name) +
  scale_edge_color_manual(values = c("lightgrey", "lightpink")) +
  scale_color_manual(values = c("grey", "lightpink")) +
  theme_graph() +
  theme(legend.position="none")
```

B is NOT pivotal for the pair consisting of D and E because because there is another shortest path which does not contain B (B is not on every shortest path between D and E). 
```{r}
# add edge data for NOT pivotal nodes
E(g1)$pivotalB <- c("Yes", "Yes", "Yes", "No","No", "No", "NotPath")

ggraph(g1, layout = "kk") + 
  geom_edge_link(aes(colour = E(g1)$pivotalB)) + 
  geom_node_point(colour = "grey", size = 7) +
  geom_node_text(label = V(g1)$name) +
  scale_edge_color_manual(values = c("gold", "lightgrey", "lightpink")) +
  theme_graph() +
  theme(legend.position="none")
```

Also note D is not pivatol for *any* pairs. 
```{r}
# D is NOT pivotal
V(g1)$pivotalD <- c("No","No","No","Yes","No","No")

ggraph(g1, layout = "kk") + 
  geom_edge_link(colour = "lightgrey") + 
  geom_node_point(aes(colour = V(g1)$pivotalD), size = 7) +
  geom_node_text(label = V(g1)$name) +
  scale_color_manual(values = c("grey", "lightpink")) +
  theme_graph() +
  theme(legend.position="none")
```

# GATEKEEPER

```{r}
g2 <- graph( c("E", "F", "F", "A", "A", "E", "A", "B", "B", "D", "D", "C", "C", "A", "D", "A"), directed = FALSE)

# When the edge list has vertex names, the number of nodes is not needed

ggraph(g2, layout = "kk") + 
  geom_edge_link(colour = "lightgrey") + 
  geom_node_point(colour = "grey", size = 7) +
  geom_node_text(label = V(g2)$name) +
  theme_graph() +
  theme(legend.position="none")
```

A node is a gatekeeper if in order to connect two neighboring nodes the edges must pass through that node. For example: node A below is a gatekeeper because it lies on every path from B to E. 

```{r}
V(g2)$globalA <- c(1,0,1,1,0,0)
E(g2)$globalA <- c(0,0,1,1,0,0,0,0)

# When the edge list has vertex names, the number of nodes is not needed

ggraph(g2, layout = "kk") + 
  geom_edge_link(aes(colour = as.factor(E(g2)$globalA))) + 
  geom_node_point(aes(colour = as.factor(V(g2)$globalA)), size = 7) +
  geom_node_text(label = V(g2)$name) +
  scale_color_manual(values = c("lightgrey", "lightblue")) +
  scale_edge_color_manual(values = c("lightgrey", "lightblue")) +
  theme_graph() +
  theme(legend.position="none")
```

The gatekeeper role can be relative to the scale at which you're looking at the graph. For instance node D is a *local* gatekeeper because it connects nodes C and B

```{r}
V(g2)$local <- c("A", "A", "A", "B", "C", "B")
E(g2)$local <- c(0,0,0,0,1,1,0,0)

# When the edge list has vertex names, the number of nodes is not needed

ggraph(g2, layout = "kk") + 
  geom_edge_link(aes(colour = as.factor(E(g2)$local))) + 
  geom_node_point(aes(colour = V(g2)$local), size = 7) +
  geom_node_text(label = V(g2)$name) +
  scale_color_manual(values = c("lightgrey", "lightblue", "lightpink")) +
  scale_edge_color_manual(values = c("lightgrey", "lightpink")) +
  theme_graph() +
  theme(legend.position="none")
```

But it is not a global gatekeeper because C and B can also be connected through node A. 

```{r}
V(g2)$global <- c(0,0,1,1,2,1)
E(g2)$global <- c(0,0,0,1,0,0,1,0)

ggraph(g2, layout = "kk") + 
  geom_edge_link(aes(colour = as.factor(E(g2)$global))) + 
  geom_node_point(aes(colour = as.factor(V(g2)$global)), size = 7) +
  geom_node_text(label = V(g2)$name) +
  scale_color_manual(values = c("lightgrey", "lightblue", "lightpink")) +
  scale_edge_color_manual(values = c("lightgrey", "lightblue")) +
  theme_graph() +
  theme(legend.position="none")
```

